Functors, CPS and monads, or how to generate efficient code from abstract designs

### 2005-04-07 paper/metamonads.tex
abstract

\begin{abstract}
We use monads and Ocaml's advanced module system to
demonstrate that, in a generative context, it is
possible to totally eliminate the overhead of abstraction.
This lets one use extreme forms of \textit{information hiding}
at no run-time cost.  Furthermore the typed nature of
the generative context provide static guarantees about the
generated code.  The various extensions
(aspects) can be made completely orthogonal and compositional,
even in the presence of name-generation for temporaries and
other bindings and ``interleaving'' of aspects.  We also
show how to encode some domain-specific knowledge so that
``clearly wrong'' compositions can be statically rejected by the
compiler.
\end{abstract}

>>>>

\begin{abstract}
With Gaussian Elimination as a representative family of numerical
algorithms, we use multi-stage programming, monads and Ocaml's
advanced module system to demonstrate the complete elimination of the
abstraction overhead while avoiding any inspection of the generated
code.  We parameterize our Gaussian Elimination code to great extent
(over domain, matrix representations, determinant tracking and
pivoting policies, result types, etc) at no run-time cost.  Because
the resulting code is generated just right and not changed afterwards,
we enjoy MetaOCaml guarantees that the generated code is well-typed.
We further demonstrate that various abstraction parameters (aspects)
can be made orthogonal and compositional, even in the presence of
name-generation for temporaries and other bindings and
``interleaving'' of aspects.  We also show how to encode some
domain-specific knowledge so that ``clearly wrong'' compositions can
be statically rejected by the compiler when processing the generator
rather than the generated code.

We use monads and Ocaml's advanced module system to
demonstrate that, in a generative context, it is
possible to totally eliminate the overhead of abstraction.
This lets one use extreme forms of \textit{information hiding}
at no run-time cost.  Furthermore the typed nature of
the generative context provide static guarantees about the
generated code.  The various extensions
(aspects) can be made completely orthogonal and compositional,
even in the presence of name-generation for temporaries and
other bindings and ``interleaving'' of aspects.  We also
show how to encode some domain-specific knowledge so that
``clearly wrong'' compositions can be statically rejected by the
compiler.
\end{abstract}
